Exploring Anomalies in Pacific Northwest Special Education Trends

### 2020-12-07 scripts/Final_project.Rmd
Removed Citation confusion and Correllogram

```{correllogram}

#cleaning data
bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))

@@

The above correlogram isolates a single school district - Greater Victoria - as a representation of the trends observed within the whole data set. This series of plots gives a snapshot introduction for the data trends to be observed.

@@

The three datasets were used for analysis in R (cite R) for reproducibility and data visualization. All code for analysis and data representation is publicly available for reproducibility (https://github.com/JoeSwinehart/datascienceproject). There are several caveats for data use and cleaning in the current analysis. First, the BC and OR datasets were not a completely parallel comparison, because the OR data set includes a broader age range of students (ages 3 - 21) as compared to the BC data set (ages 5 - 19). The Oregon data set was pared down to include a similar age range to BC data; however, it is worth noting that differences in age range may represent different approaches to diagnostic categorization (e.g., intellectual disability in BC can be diagnosed at age 5; whereas the Oregon data follows IDEA where it is more likely young children will be categorized as developmentally delayed). Second, data masking necessitates a decision-point for analysis. Specifically, the BC Ministry of Education data masks all cells with headcounts below 10 students, therefore our analyses left out these cases when computing effect sizes for provincial-level data. Data masking was most prominently an issue when examining school district-level data in BC, as smaller school districts often masked low-incidence exceptionality data. When appropriate, masked data is assigned a conservative estimate of (n = 9) such that enough data units were included in effect size comparisons. Third, Canadian census data from 2016 is the basis of classification for district geographical location (e.g., urban, suburban, rural, exurban). Due to the geographical distribution of population in BC, only two categories are used for the current analyses: urban and rural. Fourth, due to low population numbers with independent schools in BC, the current study will only examine public school data for the years 2002 - 2019.

## Analysis Plan
The majority of the analyses for the current study are descriptive in nature. The descriptive analysis includes a table or visual representation of prevalence rates or headcounts of students with exceptionalities over time. As the current study is examining anomalies in trend data, trends that break the longitudinal trendline (e.g., sudden drop in prevalence after a decade of linear increase in prevalence) may be examined further. That is, effect sizes between two groups (e.g., group one is three years before a policy change, and group two is three years after a policy change) may be examined to detect the magnitude of differences between groups if they are hypothesized to represent a change in practice stemming from policy implementation. More specifically on the calculation of effect sizes, the Hedge’s g statistic will be used because of the robustness against unequal group sizes.

>>>>

```{correllogram}
#we know this doesn't show; couldn't get it to work by submission deadline.
#cleaning data
bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))

@@

@@

The three datasets were used for analysis in R for reproducibility and data visualization. All code for analysis and data representation is publicly available for reproducibility (https://github.com/JoeSwinehart/datascienceproject). There are several caveats for data use and cleaning in the current analysis. First, the BC and OR datasets were not a completely parallel comparison, because the OR data set includes a broader age range of students (ages 3 - 21) as compared to the BC data set (ages 5 - 19). The Oregon data set was pared down to include a similar age range to BC data; however, it is worth noting that differences in age range may represent different approaches to diagnostic categorization (e.g., intellectual disability in BC can be diagnosed at age 5; whereas the Oregon data follows IDEA where it is more likely young children will be categorized as developmentally delayed). Second, data masking necessitates a decision-point for analysis. Specifically, the BC Ministry of Education data masks all cells with headcounts below 10 students, therefore our analyses left out these cases when computing effect sizes for provincial-level data. Data masking was most prominently an issue when examining school district-level data in BC, as smaller school districts often masked low-incidence exceptionality data. When appropriate, masked data is assigned a conservative estimate of (n = 9) such that enough data units were included in effect size comparisons. Third, Canadian census data from 2016 is the basis of classification for district geographical location (e.g., urban, suburban, rural, exurban). Due to the geographical distribution of population in BC, only two categories are used for the current analyses: urban and rural. Fourth, due to low population numbers with independent schools in BC, the current study will only examine public school data for the years 2002 - 2019.

## Analysis Plan
The majority of the analyses for the current study are descriptive in nature. The descriptive analysis includes a table or visual representation of prevalence rates or headcounts of students with exceptionalities over time. As the current study is examining anomalies in trend data, trends that break the longitudinal trendline (e.g., sudden drop in prevalence after a decade of linear increase in prevalence) may be examined further. That is, effect sizes between two groups (e.g., group one is three years before a policy change, and group two is three years after a policy change) may be examined to detect the magnitude of differences between groups if they are hypothesized to represent a change in practice stemming from policy implementation. More specifically on the calculation of effect sizes, the Hedges' g statistic will be used because of the robustness against unequal group sizes.
